---
title: "Performance of WHO algorithms -- Shehia level"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
   # theme: paper
    source_code: embed
#runtime: shiny
---

```{r setup, include=FALSE}


library(stringr)
library(flexdashboard)
library(leaflet)  # interative mapping
library(DT)  # interactive tables
library(crosstalk)  # inter-widget interactivity
library(tidyverse)  # tidy data manipulation

znz_sf <- sf::st_read("Zanzibar Shehia_Refactored_11012021.shp", quiet = TRUE) %>% 
   linelist::clean_data(guess_dates = FALSE) %>%  
    mutate(district = word(region_nam, 1, sep = "_")) %>% 
#   mutate(District = str_to_title(str_replace_all(district_n, "_", " ")), 
#          Shehia = str_to_title(str_replace_all(shehia_nam, "_", " "))) %>% 
  sf::st_make_valid() %>% 
  sf::st_cast("POLYGON") %>% 
  sf::st_centroid() %>% 
  select(district_n, shehia = shehia_nam) %>% 
  sf::st_cast("POINT")

coords <- sf::st_coordinates(znz_sf) %>% 
  as.data.frame()
 
znz_sf1 <- sf::st_drop_geometry(znz_sf) %>% 
  bind_cols(coords)

dash_data_a <- rio::import_list("shehia_metrics_all.xlsx") %>% 
  list_rbind(names_to = "algo") %>% 
 # pivot_wider(names_from = algo, values_from = c(4:17)) %>% 
  mutate(id = paste0(district, "_", shehia, "_", algo, "_all_all"), 
         classification = "All", 
         evaluation_set = "All")  

dash_data_b <- rio::import_list("shehia_metrics_all_local.xlsx") %>% 
  list_rbind(names_to = "algo") %>% 
 # pivot_wider(names_from = algo, values_from = c(4:17)) %>% 
  mutate(id = paste0(district, "_", shehia, "_", algo, "_local_all"), 
         classification = "Local", 
         evaluation_set = "All") 

dash_data_c <- rio::import_list("shehia_metrics_all_reporting_subset.xlsx") %>% 
  list_rbind(names_to = "algo") %>% 
 # pivot_wider(names_from = algo, values_from = c(4:17)) %>% 
  mutate(id = paste0(district, "_", shehia, "_", algo,"_all_subset"), 
         classification = "All", 
         evaluation_set = "Subset")  

dash_data_d <- rio::import_list("shehia_metrics_all_reporting_subset_local.xlsx") %>% 
  list_rbind(names_to = "algo") %>% 
 # pivot_wider(names_from = algo, values_from = c(4:17)) %>% 
  mutate(id = paste0(district, "_", shehia, "_", algo, "local_subset"), 
         classification = "Local", 
         evaluation_set = "Subset") 

data_island <- rio::import("deduped_unfiltered_data.xlsx") %>% 
  select(island, facility_district, district_for_shehia) %>% 
  count(island, facility_district, district_for_shehia)


dash_data <- bind_rows(dash_data_a, dash_data_b, dash_data_c, dash_data_d) %>% 
  select(-c(weeks, no_call, no_call_outbreak, num_outbreak)) %>% 
  mutate(island = case_when(district %in% c("chakechake", 
                                            "micheweni",
                                            "mkoani",
                                            "wete") ~ "Pemba", 
                            district %in% c("kaskazini_a",
                                            "kaskazini_b", 
                                            "kati", 
                                            "kusini", 
                                            "magharibi_a", 
                                            "magharibi_b", 
                                            "mjini") ~ "Unguja",
                            .default = "Error"), 
         algo = case_when(algo == "who_2sd" ~ "Mean + 2 SD", 
                          algo == "who_75" ~ "75th percentile", 
                          algo == "who_csum" ~ "C-SUM", 
                          .default = "error"),
           across(c(2,3), ~str_to_title(str_replace_all(.x, "_", " ")))) %>% 
  relocate(algo, .after = last_col())

#%>% 
#  select(-c(Weeks, Id)) %>% 
#  left_join(znz_sf, by = c("District", "Shehia")) %>% 
#  mutate(id = paste0(District, "_", Shehia)) %>% 
#  sf::st_as_sf() %>% 
#  sf::st_make_valid()
#

data_share <- SharedData$new(dash_data, key = ~id, group = "group1")
data_share_2sd <- SharedData$new(dash_data[dash_data$algo == "Mean + 2 SD",], 
                                   key = ~id, group = "group1")
data_share_75 <- SharedData$new(dash_data[dash_data$algo == "75th percentile",], 
                                  key = ~id, group = "group1")
data_share_csum <- SharedData$new(dash_data[dash_data$algo == "C-SUM",], 
                                    key = ~id, group = "group1")



```

Interactives {data-icon="ion-stats-bars"}
=====================================  
Inputs {.sidebar}
-------------------------------------
### Case Classification and Geography

```{r }



filter_select(id = "local", 
              label = "Case classification", 
              sharedData = data_share, 
              multiple = FALSE,
              ~classification)

filter_select(id = "evaluation_set", 
              label = "WHO Algorithm input", 
              sharedData = data_share, 
              multiple = FALSE,
              ~evaluation_set)

filter_select(id = "island", 
              label = "Island", 
              sharedData = data_share, 
              multiple = FALSE,
              ~island)

filter_select(id = "district",
              label = "District",
              sharedData = data_share,
              ~district)


filter_slider(id = "prob_ever",
              label = "Outbreak Detected Ever (%)",
              sharedData = data_share,
              ~prob_ever_detected)

filter_slider(id = "prob_first",
              label = "Outbreaks Detected First Week (%)",
              sharedData = data_share,
              ~prob_first_week)


filter_slider(id = "sensitivity_epi",
              label = "Sensitivity - Epiweek (%)",
              sharedData = data_share,
              ~sens)

filter_slider(id = "false_pos",
              label = "False Positive - Epiweek (%)",
              sharedData = data_share,
              ~false_pos)

filter_slider(id = "ppv",
              label = "Positive Predictive Value - Epiweek (%)",
              sharedData = data_share,
              ~ppv)



```

```{js}
function remove_all_option() {
    document.getElementById("local").getElementsByClassName("selectized")[0].selectize.removeOption("");
}

function remove_all_option_district() {
    document.getElementById("district").getElementsByClassName("selectized")[0].selectize.removeOption("");
}

function remove_all_option_eval() {
    document.getElementById("evaluation_set").getElementsByClassName("selectized")[0].selectize.removeOption("");
}

function filter_default() {
    document.getElementById("local").getElementsByClassName("selectized") 
[0].selectize.setValue("All", false);
 }
 
 function filter_default_eval() {
    document.getElementById("evaluation_set").getElementsByClassName("selectized") 
[0].selectize.setValue("All", false);
 }
 
 
 
 
 
 $(document).ready(function () {
  filter_default();
  filter_default_eval();
  remove_all_option();
  remove_all_option_district();
  remove_all_option_eval();
});


```


Row 
-------------------------------------
### Number of Shehia Meeting criteria
```{r}


 plotly::plot_ly(data = data_share, 
                 x = ~ algo, 
                 type = "histogram")%>% 
  plotly::layout(xaxis = list(title = 'WHO Algorithm'))
 
```

### Mean + 2 SD
```{r }

a <- 1:15
keep <- c(1:11)

b <- data_share_2sd %>% 
  datatable(rownames = FALSE,
                 colnames = c('District', 'Shehia', 'Epiweeks Outbreak (n)',
                              'Epiweeks Routine (n)', 'Number Outbreaks (n)', 
                              'False Positive Epiweeks (%)', 
                              'Epiweek Sensitiviy (%)', 'Epiweek Specificity (%)',
                              'Epiweek PPV (%)', 'Epiweek NPV (%)',
                              'Outbreaks Detected on First Week (%)',
                              'Outbreaks Ever Detected (%)'),
                    options = list(
                      columnDefs = list(list(visible=FALSE, targets = a[! a %in% keep])),
                      dom = 'ritp'))
                    

b
```




Row 
-------------------------------------
### 75th percentile
```{r}

 

b <-  data_share_75 %>% 
   datatable(rownames = FALSE,
                 colnames = c('District', 'Shehia', 'Epiweeks Outbreak',
                              'Epiweeks Routine', 'Number Outbreaks', 
                              'False Positive Epiweeks', 
                              'Epiweek Sensitiviy', 'Epiweek Specificity',
                              'Epiweek PPV', 'Epiweek NPV',
                              'Outbreaks Detected Week 1',
                              'Outbreaks Ever Detected'),
                   # extensions = 'Buttons',
                    options = list(
                      columnDefs = list(list(visible=FALSE, targets = a[! a %in% keep])),
                      dom = 'ritp'
                    #  buttons = "excel"))
                    ))

b
```

### C-SUM
```{r}


data_share_csum %>% 
   DT::datatable(rownames = FALSE,
                 colnames = c('District', 'Shehia', 'Epiweeks Outbreak (n)',
                              'Epiweeks Routine (n)', 'Number Outbreaks (n)', 
                              'False Positive Epiweeks (%)', 
                              'Epiweek Sensitiviy (%)', 'Epiweek Specificity (%)',
                              'Epiweek PPV (%)', 'Epiweek NPV (%)',
                              'Outbreaks Detected on First Week (%)',
                              'Outbreaks Ever Detected (%)'),
                    options = list(
                      columnDefs = list(list(visible=FALSE, targets = a[! a %in% keep])),
                      dom = 'ritp'))

#   datatable(filter = 'top',
#                     extensions = 'Buttons',
#                     options = list(
#                                 paging = TRUE,
#                                 searching = TRUE,
#                                 fixedColumns = TRUE,
#                                 autoWidth = TRUE,
#                                 ordering = TRUE,
#                                 dom = 'tB',
#                                 buttons = c('excel'))
#                     )
```


Data download {data-orientation=rows data-icon="fa-file-arrow-down"}
===================================== 
```{r}
data_share %>% 
  datatable(rownames = FALSE, 
            colnames = c('District', 'Shehia', 
                         'Epiweeks Outbreak (n)',
                         'Epiweeks Routine (n)',
                         'Number Outbreaks (n)', 
                              'False Positive Epiweeks (%)', 
                              'Epiweek Sensitiviy (%)', 'Epiweek Specificity (%)',
                              'Epiweek PPV (%)', 'Epiweek NPV (%)',
                              'Outbreaks Detected on First Week (%)',
                              'Outbreaks Ever Detected (%)', 
                         'Case Classification*',
                         'WHO Input',
                         'Island', 
                         'WHO Algorithm'),
            extensions = 'Buttons',
                    options = list(
                      columnDefs = list(list(visible=FALSE, targets = c(12))),
                      dom = 'Britp',
                      buttons = "excel")
  )
```

How to use {data-orientation=rows data-icon="fa-info-circle"}
===================================== 

The intent of this dashboard is to provide insight into which shehias may be 
viable options for the implementation of WHO-recommended algorithms for outbreak
monitoring. We understand that the goal of the program is to implement outbreak
detection schemes at the lowest organizational level possible and agree with this
approach as malaria outbreaks can be very focal. However, we also understand
that algorthim-based approaches are not feasible or appropriate for all shehias
in Zanzibar, as the low malaria burden and limited personnel/resources in some 
areas require a more practical approach. 


Users may simultaneously apply multiple filters across the dataset to determine
how many, and the identity of the shehias in which different algorithms are 
performing well, or alternatively not well.

### Interactives
#### Inputs
The bar on the left-hand side of the 'Interactives' tab allows users to enter 
characteristics of interest. 'Case classification' allows the the user to examine
the performance of WHO-recommended algorithms when _either_ all reported malaria
cases are consider, or only cases that were not determined to be imported are included. 
Note that in the case-based database, there are several designations with the 
following distribution:

* Imported - 19806
* In progress - 3899
* Indigenous - 37515
* Induced - 196
* Introduced - 666
* Lost to followup - 2015
* Relapse - 54
* Relapsing - 20
* Unclassified - 1073

For the purposes of this classification 'Local' are those that are classified as
anything _other_ than imported and introduced. 

'WHO Algorithm Input' allows the user to determine which dataset is used by the 
WHO-recommended algorithms. Part of this exercise involved censoring cases in 
the routine data that appeared to come from facilities that did not consistently
report cases over time. As reporting bias can have a large effect on trends in 
case counts, only cases from consistently reporting facilities were used in the 
gold-standard definition of outbreaks. By default, assessment of WHO algorithms
was done by inputting _all_ available case data as this is more likely to be a
reflection of true performance without an extensive data cleaning effort. However,
this option allows the user to determine how WHO algorithms would perform after
benefiting from a rigorous data cleaning, a type of "best-case" scenario.

'Island' and 'District' filters reflect the geographic characteristics of the 
case; 'District' allows selection of multiple districts.

The slider filters allow the user to selectively view shehias by the performance
of WHO-recommended outbreak detection algorithms there. Five performance-based
filters are available, all of which are displayed as a percentage. Numerators and
denominators for each are listed below: 

* Outbreak Detected Ever - Number of true outbreaks occurring in the shehia for 
which the WHO-recommended algorithm detected _at least_ one anomalous epiweek
during its duration divided by number of distinct true outbreaks occurring in 
the shehia

* Outbreak Detected First Week - Number of true outbreaks occurring in the shehia for 
which the WHO-recommended algorithm detected an anomalous epiweek on
_its first week_ divided by number of distinct true outbreaks occurring in 
the shehia.

* Sensitivity Epiweek - The number of epiweeks in the shehia that were both
part of a true outbreak and detected by the WHO-algorithm divided by the number
of epiweeks in the shehia that were part of a true outbreak.

* False Positive Epiweek - The number of epiweeks in the shehia that were both
flagged by the WHO-algorithm as an anomaly but were _not_ part of a true 
outbreak divided by the number of epiweeks in the shehia that were _not_ part 
of a true outbreak.

* Positive Predictive Value Epiweek - The number of epiweeks in the shehia that 
were both part of a true outbreak and detected by the WHO-algorithm divided by the number
of epiweeks that were flagged by the WHO-algorithm as anomalous.

#### Outputs
The bar chart in the upper left panel shows the number of shehias that match the 
criteria entered by the user in the left-hand panel. Though this graph was made
with coding that allows for interactivity, we recommend not attempting to click 
or filter using this graph as this does not add functionality to the tables and
will likely only present a confusing display.

Each table displays the shehias matching the criteria in the left-hand filter
panel. A count of the entries included in each is present just below its title;
however, note that the number of total entries is provided parenthetically and
that this number represents the total number of rows present in the overall 
dataset _not_ the number of shehias for which there are data. Each shehia has 
_multiple_ rows to corresponding the different scenarios available (local cases 
vs. all cases, data from consistent reporters only vs. all data, etc.). 
Definitions of columns are identical to those above with some additions.
'Epiweeks outbreak' shows the number of epiweeks in each shehia when a true 
outbreak was ongoing according to the gold-standard definition in this exercise'
'Epiweeks routine' displays the number of epiweeks in each shehia that did not 
meet the definition of an outbreak provided in other files. 

### Data Download
All data matching the criteria entered in the filter sidebar are displayed here.
Note that each row corresponds to a location and algorithm _combination_, so it
is possible for each shehia to appear as many as three times. 

The intended purpose of the table is to enable the user to download the data of
interest (expressed through the filters) for comparison and deeper analysis. 
Clicking the grey button labeled 'Excel' in the upper left-hand corner will open
a prompt allowing the data to be saved as an Excel spreadsheet. Note that the 
default filename provided will be "Performance of WHO algorithms -- Shehia level.xlsx".
If more than one such file with different filters is applied is downloaded, it is
recommended to provide additional information in the filename. At the very least, 
the names must be changed with subsequent downloads, or the files of interest will
be overwritten. 

Consistent with above, one additional column labeled 'id' that is not shown in 
the table will be present in the xlsx file. These data are simply a concatenation 
of district, shehia, case classification, and WHO algorithm, which served as a
handle for creation of the linked outputs on the 'Interactives' tab.

